Men and past smokers from ontario are potentially more physically active than women and non-smokers

### 2024-12-04 paper/paper.qmd
Change of map chunk for paper.qmd
Changed script from 08-health_region.R to 05-health_region.R

---
title: "Men and past smokers from ontario are potentially more physically active than women and non-smokers respectively"
subtitle: "Women are engaging in lesser physical activities than men. Furthermore, individuals who smoked before are more likely to engage in physical activity. Consumption of stimulant-related drugs could have also increased as well, resulting in increased overall physical activities in Ontario."
author: 
  - Chris Yong Hong Sen
thanks: "Code and data are available at: [https://github.com/Monoji77/physical_health_ontario](https://github.com/Monoji77/physical_health_ontario)."
date: today
date-format: long
abstract: "This study explores factors influencing physical activity levels in Ontario, focusing on demographic and substance abuse. Our findings reveal that women engage in less physical activity than men, while individuals who have smoked in the past and those consuming stimulant-related drugs are more likely to be physically active. Additionally, age, highest educational attainment, and health region also play significant roles in determining physical activity levels. These findings highlight the complex interplay of gender, lifestyle behaviors, and sociodemographic factors, providing helpful information for designing targeted public health interventions to promote physical activity."
format: 
  pdf:
    fig-pos: H
number-sections: true
bibliography: references.bib
---

@@

Physical activity is a cornerstone of public health, yet participation rates remain suboptimal across various populations in Ontario. This is especially concerning as physical inactivity is a leading contributor to chronic diseases such as heart disease, diabetes, and mental health disorders. Factors such as age, sex, educational attainment, alcohol consumption, smoking habits, and regional disparities in health access play an important role in shaping physical activity behaviors. Understanding these factors and their complex interactions is important for developing targeted interventions that can improve health outcomes and reduce health inequalities in Ontario.

@@

The remainder of this paper is structured as follows. @sec-data would provide a sense of the chosen Computing in the Humanities and Social Sciences dataset as well as data exploration of interested variables. @sec-Model goes into the details of the linear regression model fitted on the predictors of interest variables, for which it's results would be showcased in @sec-results. Finally @sec-discussion would interpret the results of the fitted linear regression model, which includes limitations of the paper. At the end of the paper in [Appendix -@sec-diagnostic_checks], this contains diagnostic checks for all models considered, and in [Appendix -@sec-survey_methodology], it contains a section regarding the survey and it's methodology.

Overall, this paper was generated using R programming language [@citeR]. Reading of raw data and data manipulation was performed using `Tidyverse` package [@citeTidyverse] and parquet file handling was done using functions in `arrow` package [@citeArrow]. Shape date was read into R using `sf` package, and generation of maps were done using functions in `tmap` package [@citeSF; @citationTmap]. Any other visual plots were using `ggplot`-related functions in Tidyverse [@citeTidyverse] and tables were generated using `kable` function in `knitr` package [@citeKnitR].

@@

The dataset chosen is obtained from Computing in the Humanities and Social Sciences/CHASS at the university of Toronto [@SDA_CHASS], specifically the Canadian Community Health Survey (CCHS) annual component 2017-2018 which was collected by @StatisticsOfCanada. The goal of the CCHS survey is to collect reliable health-related data on the Canadian population to support health surveillance, research, and program evaluation. It aims to provide timely, accessible, and flexible data for monitoring health trends, studying small populations and rare characteristics, and addressing emerging health issues, ultimately helping improve the health and well-being of Canadians.

@@

According to @fig-barplot-sex, there is roughly an equal proportion of sexes from our dataset, with slightly more males (6082 respondents) than females (5859 respondents). Since this was a health survey conducted over phone call, there is an equal chance of a respondent to be male or female, which is roughly reflected in @fig-barplot-sex.
```{r sex predictor barplot}
#| echo: false
#| warning: false
#| message: false
#| label: fig-barplot-sex
#| fig-cap: Barplot of sex
data_cleaned_1 |>
  mutate(sex = factor(sex, levels = c(0,1), labels =c('Female','Male'))) |>
  ggplot()  +
  geom_bar(aes(x=sex, fill=sex), show.legend=F) +
  scale_fill_manual(values = c('Female' = 'indianred4',
                               'Male' = 'skyblue1')) +
  theme_minimal() +
  xlab('Individual\'s sex') +
  ylab('Number of individuals') +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

@@

Overall, since the residual assumptions in linear regression are satisfied, we are able to use the coefficient estimates obtained from our final fitted linear model to gain understanding in the relationship between our predictors and the response as discussed in @sec-discussion.

>>>>

---
title: "Men and past smokers from ontario are potentially more physically active than women and non-smokers"
subtitle: "Women are engaging in lesser physical activities than men. Furthermore, individuals who smoked before are more likely to engage in physical activity. Consumption of stimulant-related drugs could have also increased as well, resulting in increased overall physical activities in Ontario."
author: 
  - Chris Yong Hong Sen
thanks: "Code and data are available at: [https://github.com/Monoji77/Alcohol_Use_Ontario](https://github.com/Monoji77/Alcohol_Use_Ontario)."
date: today
date-format: long
abstract: "This study explores factors influencing physical activity levels in Ontario, focusing on demographic and substance abuse. Our findings reveal that women engage in less physical activity than men, while individuals who have smoked in the past and those consuming stimulant-related drugs are more likely to be physically active. Additionally, age, highest educational attainment, and health region also play significant roles in determining physical activity levels. These insights highlight the complex interplay of gender, lifestyle behaviors, and sociodemographic factors, providing valuable information for designing targeted public health interventions to promote physical activity."
format: 
  pdf:
    fig-pos: H
number-sections: true
bibliography: references.bib
---

@@

Physical activity is a cornerstone of public health, yet participation rates remain suboptimal across various populations in Ontario. This is especially concerning as physical inactivity is a leading contributor to chronic diseases such as heart disease, diabetes, and mental health disorders. Factors such as age, sex, educational attainment, alcohol consumption, smoking habits, and regional disparities in health access play a critical role in shaping physical activity behaviors. Understanding these factors and their complex interactions is crucial for developing targeted interventions that can improve health outcomes and reduce health inequalities in Ontario.

@@

The remainder of this paper is structured as follows. @sec-data would provide a sense of the chosen Computing in the Humanities and Social Sciences dataset as well as data exploration of interested variables. @sec-Model goes into the details of the linear regression model fitted on the predictors of interest variables, for which it's results would be showcased in @sec-results. Finally @sec-discussion would interpret the results of the fitted linear regression model, which includes limitations of the paper. At the end of the paper in [Appendix -@sec-diagnostic_checks], this contains diagnostic checks for all models considered.

Overall, this paper was generated using R programming language [@citeR]. Reading of raw data and data manipulation was performed using `Tidyverse` package [@citeTidyverse] and parquet file handling was done using functions in `arrow` package [@citeArrow]. Generation of maps were done using functions in `tmap` package [@citationTmap]. Any other visual plots were using `ggplot`-related functions in Tidyverse [@citeTidyverse] and tables were generated using `kable` function in `knitr` package [@citeKnitR].

@@

The dataset chosen is obtained from Computing in the Humanities and Social Sciences/CHASS at the university of Toronto [@SDA_CHASS], specifically the Canadian Community Health Survey (CCHS) annual component 2017-2018 which was collected by @StatisticsOfCanada. The goal of the CCHS survey is to collect comprehensive and reliable health-related data on the Canadian population to support health surveillance, research, and program evaluation. It aims to provide timely, accessible, and flexible data for monitoring health trends, studying small populations and rare characteristics, and addressing emerging health issues, ultimately helping improve the health and well-being of Canadians.

@@

According to @fig-barplot-sex, there is roughly an equal proportion of sexes from our dataset, with slightly more males (6082 respondents) than females (5859 respondents). Since this was a health survey conducted over phone call, there is an equal chance of a respondent to be male or female, which is roughly reflected ing @fig-barplot-sex.
```{r sex predictor barplot}
#| echo: false
#| warning: false
#| message: false
#| label: fig-barplot-sex
#| fig-cap: Barplot of sex
data_cleaned_1 |>
  mutate(sex = factor(sex, levels = c(0,1), labels =c('Female','Male'))) |>
  ggplot()  +
  geom_bar(aes(x=sex, fill=sex), show.legend=F) +
  scale_fill_manual(values = c('Female' = 'indianred4',
                               'Male' = 'skyblue1')) +
  theme_minimal() +
  xlab('Individual\'s sex') +
  ylab('Number of individuals') +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

@@

Overall, since the residual assumptions in linear regression are satisfied, we are able to use the coefficient estimates obtained from our final fitted linear model to gain insights into relationship between our predictors and the response as discussed in @sec-discussion.
